\section{An hypothetical case}

Hi. My name is John Doe, and I've just graduated from high school. I'm a bit
undecided about what to do now, what path to follow. My mom  says that an
attorney is a really respected profession, and I could earn a lot of money so I
could help her afterwards, so I'll take a shot a Law.

I'm a poor person, so I don't have any money to afford a private university. And
the public ones are so competitive that I would have to stop working and focus
for a long time on studying for the vestibular just to have a shot to be
admitted. But I've heard of a new program by the federal government that
eliminates the need for a vestibular, and gives me the liberty of studying at my
own pace and time, but it uses computers and my family doesn't have one, by the
obvious reason of having to buy food instead.

But there's a new computer center that the government opened at my community,
and they say one of their objectives is to attend to people in situations like
mine, so I'll schedule appointments on there.

\textit{six months later \ldots }

I'm studying every day for the second certification. In the first four months
I've learned enough to pass the first one with a acceptable grade. I'm planning
to study more this time to make it with a higher grade. People say that
university has plenty of vacancy without all this subjects on the classroom, but
I don't know how it will be when I finish all the certifications, so it's better
to be safe than sorry.

The materials are fun, and I've met several people who were studying the same
subjects, and we kind of help each other on the suggested lessons and
assignments. I think the next test will be hard, but with each passing day I'm
learning more to communicate my questions on-line and I think I'll make it
through.

\textit{a couple of years later \ldots }

I've made it! I've went through all the necessary certifications with a
satisfactory enough grade, and now I have been called into a model forum on a
federal facility for students! I feel ready for it, although a bit nervous, 
but at least this forum is on my town, and perhaps I could help my own community
through my work and knowledge.

\textit{a year and a half later \ldots }

I'm almost finished with my work at the forum. I've met several students and
professors, but the most pleasurable experience was to help families similar to
mine on the process. 

I've applied my knowledge directly on the field, and frankly, my certifications
did come in handy, because I had a intellectual security to handle a lot of
different situations, and the ones I didn't knew how to handle I used the same
research process I've used back on my home study days to figure it out. In some
months my graduation will take place, I'll miss this my colleagues, but I'm also
anxious to get on the market and flourish as a professional...



